---
title: "Exploratory Analysis"
output:
  html_document:
    df_print: paged
---

Exploratory analysis before fitting regression models


# Process Data
```{r processing, message=FALSE}
library(dplyr);
library(rlang);

delays <- read.csv('data/delays.csv');

# add timestamp
delays <- delays %>% 
    mutate(
        timestamp = as.POSIXct(
            paste(Date, Time), 
            format = '%Y/%m/%d %H:%M'
            )
        );
```


Let's get a summary of each delay code. 

```{r codes}
delay.summary <- delays %>% 
    group_by(Code) %>%
    summarize(
        count = n(), 
        mean.delay = mean(Min.Delay), 
        min.delay = min(Min.Delay),
        median.delay = median(Min.Delay),
        max.delay = max(Min.Delay)
        ) %>%
    ungroup() %>%
    filter(max.delay > 0) %>%
    arrange(-count);
```

Plot the histogram of all delays.

```{r histograms}
group.histogram <- function(input.data, filter.column = NULL, value = NULL) {

    if( !is.null(filter.column) ) {
       group.data <- input.data %>%
           filter( !!sym(filter.column) == value); 
    } else {
        group.data <- input.data;
    }
    
    graphics::hist(
        group.data$Min.Delay, 
        main = value,
        xlab = 'Delay (minutes)',
        ylab = 'Freq',
        col = 'steelblue',
        xaxs = 'i',
        yaxs = 'i'
        );
}

group.histogram(delays);
```

There are some extreme delays that make the histogram impossible to work with. Let's have a look.

```{r}
delays %>% filter(Min.Delay >= 100);
```

The histogram is more legible if we restrict it to delays under an hour.

```{r}
group.histogram(delays %>% filter(Min.Delay < 60) );
```

Now we can break it down by code.
```{r, fig.width=9, fig.height=7}
short.delays <- delays %>% filter(Min.Delay < 60);

par(mfrow = c(3, 4));
for( code in delay.summary$Code[1:12] ) group.histogram(short.delays, 'Code', code);
```

Or station
```{r, fig.width=9, fig.height=7}
delay.summary <- delays %>% 
    group_by(Station) %>%
    summarize(
        count = n(), 
        mean.delay = mean(Min.Delay), 
        min.delay = min(Min.Delay),
        median.delay = median(Min.Delay),
        max.delay = max(Min.Delay)
        ) %>%
    ungroup() %>%
    filter(max.delay > 0) %>%
    arrange(-count);

par(mfrow = c(3, 4));
for( station in delay.summary$Station[1:12] ) group.histogram(short.delays, 'Station', station);
```

Interistingly, there are no clear differences in distribution between the station. By contrast, there are clear differences by type of delay.
